Ontology Matching with word2vec

### 2016-06-17 Paper/Draft.tex
Added example to OWL

 \subsubsection{The state of the art} 
 Recently, an effective and efficient skip-gram model has been developed, called \textf{Word2vec}, which uses a number of extensions over previous methods that enable it to efficiently learn an effective representation\cite{w2v}.
 \textf{Word2vec} learns this representation by trying to predict the context of a word. First the vocabulary is one-hot encoded. Then the model learns to predict a context vector that minimizes the distance to the context one-hot encoded vectors. The model is a neural network with one hidden layer. Since the input is one-hot encoded, every word in the vocabulary is represented by one row in the weight matrix between the input layer and hidden layer. This row is the vector representation of the word.

 \subsubsection{Extensions}
 \textf{Word2vec} has been expanded in a number of different ways. For example, word order can be preserved, leading to a similarity measure that is closer to syntax, as syntax defines which words go where in a sentence. This preservation was done by having multiple output layers, one for every word in the context. The research shows that the input does not have to be a bag-of-words representation, which will be useful later on\cite{ling2015}.
 Another interesting development is the multi-sense \textf{Word2vec}, which allows for multiple vector representations per word, depending on the number of different definitions a word has. For example, the word bank represents both the monetary institute and the riverside, which would both have a different representation in this extension. It can differentiate between two different meanings based on the context. This will help with resolving ambiguity, a very important factor in ontology alignment\cite{multisense}.

@@

 \newpage
 \section{Method}
 This Section will describe the method that I developed. Firstly, I will describe the general idea. In the second section, I will give a theoretical justification for why this method should work and how it works conceptually. After that I will describe the different ways that the input graphs can be used to train the \textf{Word2vec} model and create the alignment, some of which I have discarded on theoretical grounds as I will explain in the relevant section. Lastly, I will talk about the implementation process.

@@

 \textf{Word2vec} is an unsupervised method that implicitly maps words to a vector space that has been shown to contain semantic properties. Explicitly, it predicts words from the context of that word. %cite
 It can be trained on large corpora quickly and the resulting vectors can easily be used to measure the distance between words. All these properties make it an excellent candidate for a concept distance measure to use for ontology matching. As an added benefit, it can also take into account the context of a concept, which may be able to improve its performance over a matcher that only uses the label of a concept.

@@

 To solve these problems I will adapt \textf{Word2vec} to graphs to create an algorithm that converts nodes to vectors, in other words \textf{Node2vec}. It can be extended with multi-sense embeddings, different ways of considering the neighbourhood and hot-starting. %TODO this paragraph needs to be explained

@@

  The most basic \textf{Node2vec} model converts a node to a vector purely based on its own label. This method has one advantage over simple string matching: if no node is found in the other ontology that matches exactly, we can still find a node that is similar since it is close in the vector space. Therefore, this method should already be an improvement over the most basic string matching algorithms. It may even be competitive with more advanced string matchers (that look at substrings of labels). For example, \emph{ear lobe} and \emph{ear} will be matched by a substring matcher, but are also semantically similar since they will often be mentioned in the same context.

@@

 \subsection{Implementation}
 The two foundations the algorithm has been build and tested on are Factorie, described in \ref{factorie}, and AgreementMaker, described in \ref{agreementmaker}, respectively. Earlier work on a forked branch of factorie was done by \cite{multisensecode} to implement multi-sense \textf{Word2vec}.

>>>>

 \subsubsection{The state of the art} 
 Recently, an effective and efficient skip-gram model has been developed, called \textbf{Word2vec}, which uses a number of extensions over previous methods that enable it to efficiently learn an effective representation\cite{w2v}.
 \textbf{Word2vec} learns this representation by trying to predict the context of a word. First the vocabulary is one-hot encoded. Then the model learns to predict a context vector that minimizes the distance to the context one-hot encoded vectors. The model is a neural network with one hidden layer. Since the input is one-hot encoded, every word in the vocabulary is represented by one row in the weight matrix between the input layer and hidden layer. This row is the vector representation of the word.

 \subsubsection{Extensions}
 \textbf{Word2vec} has been expanded in a number of different ways. For example, word order can be preserved, leading to a similarity measure that is closer to syntax, as syntax defines which words go where in a sentence. This preservation was done by having multiple output layers, one for every word in the context. The research shows that the input does not have to be a bag-of-words representation, which will be useful later on\cite{ling2015}.
 Another interesting development is the multi-sense \textbf{Word2vec}, which allows for multiple vector representations per word, depending on the number of different definitions a word has. For example, the word bank represents both the monetary institute and the riverside, which would both have a different representation in this extension. It can differentiate between two different meanings based on the context. This will help with resolving ambiguity, a very important factor in ontology alignment\cite{multisense}.

@@

 \newpage
 \section{Method}
 This Section will describe the method that I developed. Firstly, I will describe the general idea. In the second section, I will give a theoretical justification for why this method should work and how it works conceptually. After that I will describe the different ways that the input graphs can be used to train the \textbf{Word2vec} model and create the alignment, some of which I have discarded on theoretical grounds as I will explain in the relevant section. Lastly, I will talk about the implementation process.

@@

 \textbf{Word2vec} is an unsupervised method that implicitly maps words to a vector space that has been shown to contain semantic properties. Explicitly, it predicts words from the context of that word. %cite
 It can be trained on large corpora quickly and the resulting vectors can easily be used to measure the distance between words. All these properties make it an excellent candidate for a concept distance measure to use for ontology matching. As an added benefit, it can also take into account the context of a concept, which may be able to improve its performance over a matcher that only uses the label of a concept.

@@

 To solve these problems I will adapt \textbf{Word2vec} to graphs to create an algorithm that converts nodes to vectors, in other words \textbf{Node2vec}. It can be extended with multi-sense embeddings, different ways of considering the neighbourhood and hot-starting. %TODO this paragraph needs to be explained

@@

  The most basic \textbf{Node2vec} model converts a node to a vector purely based on its own label. This method has one advantage over simple string matching: if no node is found in the other ontology that matches exactly, we can still find a node that is similar since it is close in the vector space. Therefore, this method should already be an improvement over the most basic string matching algorithms. It may even be competitive with more advanced string matchers (that look at substrings of labels). For example, \emph{ear lobe} and \emph{ear} will be matched by a substring matcher, but are also semantically similar since they will often be mentioned in the same context.

@@

 \subsection{Implementation}
 The two foundations the algorithm has been build and tested on are Factorie, described in \ref{factorie}, and AgreementMaker, described in \ref{agreementmaker}, respectively. Earlier work on a forked branch of factorie was done by \cite{multisensecode} to implement multi-sense \textbf{Word2vec}.
